### 2024-11-23 paper/paper.qmd
do data visualization

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(dplyr)
library(here)
```

```{r}
data <- read_csv(here("data", "01-analysis_data", "analysis_data.csv"))
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

```{r}

```

Talk way more about it.

```{r}
#| echo: false
#| warning: false
#| label: fig-demographics-distribution
#| fig-cap: "Demographics Distribution and Average Income by Group"
#| fig-subcap: [
#| "Gender Proportion", "Average Income by Gender", 
#| "Race Proportion", "Average Income by Race", 
#| "Region Proportion", "Average Income by Region", 
#| "Education Proportion", "Average Income by Education Level"
#| ]
#| layout-ncol: 2

# Gender proportion and income
gender_count <- data %>%
  group_by(gender) %>%
  summarise(count = n()) %>%
  mutate(percentage = round(count / sum(count) * 100, 1))

pie_chart_gender <- ggplot(gender_count, aes(x = "", y = percentage, fill = gender)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  geom_text(aes(label = paste0(percentage, "%")),
            position = position_stack(vjust = 0.5), size = 3) +
  theme_void() +
  theme(legend.position = "right") +
  labs(fill = "Gender")

avg_income_gender <- data %>%
  group_by(gender) %>%
  summarise(Average_Income = mean(INCWAGE, na.rm = TRUE))

bar_chart_gender <- ggplot(avg_income_gender, aes(x = gender, y = Average_Income, fill = gender)) +
  geom_bar(stat = "identity", width = 0.6) +
  geom_text(aes(label = round(Average_Income, 0)), 
            vjust = -0.5, size = 3) +
  labs(x = "Gender", y = "Average Income") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +  # Expand upper limit
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

# Race proportion and income
race_count <- data %>%
  group_by(race_group) %>%
  summarise(count = n()) %>%
  mutate(percentage = round(count / sum(count) * 100, 1))

pie_chart_race <- ggplot(race_count, aes(x = "", y = percentage, fill = race_group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  geom_text(aes(label = paste0(percentage, "%")),
            position = position_stack(vjust = 0.5), size = 3) +
  theme_void() +
  theme(legend.position = "right") +
  labs(fill = "Race Group")

avg_income_race <- data %>%
  group_by(race_group) %>%
  summarise(Average_Income = mean(INCWAGE, na.rm = TRUE))

bar_chart_race <- ggplot(avg_income_race, aes(x = race_group, y = Average_Income, fill = race_group)) +
  geom_bar(stat = "identity", width = 0.6) +
  geom_text(aes(label = round(Average_Income, 0)), 
            vjust = -0.5, size = 3) +
  labs(x = "Race Group", y = "Average Income") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +  # Expand upper limit
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

# Region proportion and income
region_count <- data %>%
  group_by(region) %>%
  summarise(count = n()) %>%
  mutate(percentage = round(count / sum(count) * 100, 1))

pie_chart_region <- ggplot(region_count, aes(x = "", y = percentage, fill = region)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  geom_text(aes(label = paste0(percentage, "%")),
            position = position_stack(vjust = 0.5), size = 3) +
  theme_void() +
  theme(legend.position = "right") +
  labs(fill = "Region")

avg_income_region <- data %>%
  group_by(region) %>%
  summarise(Average_Income = mean(INCWAGE, na.rm = TRUE))

bar_chart_region <- ggplot(avg_income_region, aes(x = region, y = Average_Income, fill = region)) +
  geom_bar(stat = "identity", width = 0.6) +
  geom_text(aes(label = round(Average_Income, 0)), 
            vjust = -0.5, size = 3) +
  labs(x = "Region", y = "Average Income") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +  # Expand upper limit
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )
# Education proportion and income
education_count <- data %>%
  group_by(education_level) %>%
  summarise(count = n()) %>%
  mutate(percentage = round(count / sum(count) * 100, 1))

# 设置学历顺序
education_levels <- c("Below_High_School", "High_School", "Some_College", "Bachelor", "Above_Bachelor")
education_count <- education_count %>%
  mutate(education_level = factor(education_level, levels = education_levels))

avg_income_education <- data %>%
  group_by(education_level) %>%
  summarise(Average_Income = mean(INCWAGE, na.rm = TRUE)) %>%
  mutate(education_level = factor(education_level, levels = education_levels))

# Pie chart: education level proportion
pie_chart_education <- ggplot(education_count, aes(x = "", y = percentage, fill = education_level)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  geom_text(aes(label = paste0(percentage, "%")),
            position = position_stack(vjust = 0.5), size = 3) +
  theme_void() +
  theme(legend.position = "right") +
  labs(fill = "Education Level")

# Bar chart: average income by education level, ordered
bar_chart_education <- ggplot(avg_income_education, aes(x = education_level, y = Average_Income, fill = education_level)) +
  geom_bar(stat = "identity", width = 0.6) +
  geom_text(aes(label = round(Average_Income, 0)), 
            vjust = -0.5, size = 3) +
  labs(x = "Education Level", y = "Average Income") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2)), labels = scales::comma) +  # Use comma format for y-axis
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )
# Output all eight charts in four rows with two charts per row
pie_chart_gender
bar_chart_gender
pie_chart_race
bar_chart_race
pie_chart_region
bar_chart_region
pie_chart_education
bar_chart_education

```

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

\newpage

# References